---
title: "Main Project Code"
author: "Travis Helm"
date: "11/3/2021"
output: html_document
---

# Peer Graded Assignment 1

## Loading and Processing Data (NOTE:  I saved in my folder unzipped).

I'm adding a weekend/weekdy column

```{r read_csv, echo=TRUE}
library(dplyr)

activity <- read.csv('activity.csv') 

activity <- activity %>% 
  mutate(date = as.Date(activity$date, "%Y-%m-%d"),
         Date_Type = if_else(weekdays(date) %in% c('Saturday','Sunday'), 'weekend','weekday'),
         Weekday = weekdays(date))
head(activity)
```

## What is mean total number of steps taken per day?

### Create Histogram

I group by date and then get the total number of steps

```{r Daily Steps, echo=TRUE}
library(ggplot2)
Total_Daily_Steps <- activity %>% 
  group_by(date) %>% 
  summarise(Total_Steps = sum(steps, na.rm=TRUE)) %>% 
  ungroup()

ggplot(data=Total_Daily_Steps, aes(x=Total_Steps)) +
  geom_histogram() +
  labs(x='Total Daily Steps')


```


### Summary statistics (including the mean and median) are shown below.  

```{r Mean Median, echo=TRUE}
summary(Total_Daily_Steps$Total_Steps)


```

## What is the average daily activity pattern?

I group by interval and get the total number of steps

```{r Time Series Plot}

Ave_Steps <- activity %>% 
  group_by(interval) %>% 
  summarise(Ave_Steps = mean(steps, na.rm=TRUE)) %>% 
  ungroup() 

library(lattice)
xyplot(Ave_Steps ~ interval, data = Ave_Steps, type = c("l"))


```


### What five minute interval contains the most steps?

Interval 835 (see below)

```{r Most number of steps}

max_interval <- Ave_Steps[Ave_Steps$Ave_Steps == max(Ave_Steps$Ave_Steps),]
max_interval

```

## Imputing Missing Values
There are 2304 missing cases

```{r complete cases, echo=FALSE}
  sum(!complete.cases(activity))
```

### Creating Dataset to replace missing values with average for interval

I create a column called steps_noNA that has the average steps for an interval
when the steps for that day aren't actually known

```{r Activity2}

activity2 <- activity %>% 
  left_join(Ave_Steps, by = 'interval') %>% 
  mutate(steps_noNA = ifelse(is.na(steps), Ave_Steps, steps))
```

### Histogram with corrected values

```{r histogram2}
Total_Daily_Steps2 <- activity2 %>% 
  group_by(date) %>% 
  summarise(Total_StepsNA = sum(steps, na.rm=TRUE),
            Total_StepsNoNA = sum(steps_noNA)) %>% 
  ungroup()

```

### Mean and Median with and without NAs

The mean and median are lower when missing values are present

```{r Mean and Median NA and not}

summary(Total_Daily_Steps2[,c('Total_StepsNA','Total_StepsNoNA')])
```

##Plotting by weekend/weekday

```{r weekend weekday, echo=FALSE}



Ave_Steps_Interval <- activity %>% 
  group_by(interval, Date_Type) %>% 
  summarise(Ave_Steps = mean(steps, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(Date_Type = as.factor(Date_Type))

qplot(interval, Ave_Steps, data=Ave_Steps_Interval, geom=c("line"), facets = Date_Type~., ylab = 'Number of Steps')


```
